---
title: "Lecture 11 Structural Bioinformatics"
author: "Tuan"
date: "2/15/2019"
output: github_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

## PDB data
```{r}
stats <- read.csv("Data Export Summary.csv", row.names = 1)
stats
```

Q: determine the percentage of structures solved by X-Ray and Electron Microscopy.

```{r}
pre.by.method <- stats$Total/sum(stats$Total)*100
names(pre.by.method) <- row.names(stats)
pre.by.method
```

Q:  what proportion of structures are protein?
```{r}
round(sum(stats$Proteins)/sum(stats$Total)*100, 2)
```

The **datapasta** package helps paste data in various forms as data frame form.

Section 3: Bio3D

```{r}
library(bio3d)
```


```{r}
pdb <- read.pdb("1HSG")
pdb
```

```{r}
library("bio3d.view")
```

```{r}
view(pdb, "overview", col = "sse")
```


Extract the protein only portion of this PDB structure and write out to a new PDB file.

```{r}
HIVII.protease <- atom.select(pdb, "calpha")
HIVII.protease
```
```{r}
pdb$atom[HIVII.protease$atom, ]
protein.pdb <- trim.pdb(pdb, HIVII.protease)
protein.pdb
```

```{r}
view(protein.pdb)
```

```{r}
write.pdb(protein.pdb, file = "HIVIIprotease.protein")
```


Extract the ligand (i.e. drug) and write out to a separate file.


```{r}
drug.pdb <- atom.select(pdb, "ligand")
ligand.pdb <- trim.pdb(pdb, drug.pdb)
view(ligand.pdb)
write.pdb(ligand.pdb, file = "HIVIIProtease.drug")
```


```{r}
modes <- nma(pdb)
modes <- nma(pdb)
m7 <- mktrj(modes, mode = 7, file = "mode_7.pdb")
view(m7, col = vec2color(rmsf(m7)))
```


