---
title: "Lecture 09 PCA and Clustering project"
author: "Tuan"
date: "2/8/2019"
output: github_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

## R Markdown
```{r}
fna.data <- "WisconsinCancer.csv"
wisc.df <- read.csv(fna.data)
head(wisc.df)
```

Looks like there is a strange column "X" at the end. Let's check how many samples (i.e. patents) and features (i.e. cols) are in data
```{r}
nrow(wisc.df)
ncol(wisc.df)
```

```{r}
wisc.data <- wisc.df[, 3:32]
head(wisc.data)
```

Set the row names back to the patient ID
```{r}
rownames(wisc.data) <- wisc.df$id
head(wisc.data)
```

Equivalent to these 2 code snippet is *wisc.data <- wisc.df[, c(1, 3:32)]*

Q: How many cancer/non-cancer samples
```{r}
table(wisc.df$diagnosis)
```

Q: How many features are "_mean" values?
```{r}
#colnames(wisc.data)
length(grep("_mean",colnames(wisc.data)))
```

```{r}
inds <- grep("_mean", colnames(wisc.data))
colnames(wisc.data)[inds]
```

Checking whether the data need to be scaled, because:
- Data value may be measured in different units (for an individual feature)
- Data value may have significantly different variances.

Calculate the mean of every column (second argument: 1 = rows, 2 = cols)
```{r}
round(apply(wisc.data, 2, mean), 2)
```

Similarly, for standard deviation
```{r}
round(apply(wisc.data, 2, sd), 2)
```

We need scaling, as the sd of some features are so much larger compared to others.
```{r}
# Perform PCA on wisc.data
wisc.pr <- prcomp(wisc.data, scale = TRUE)
summary(wisc.pr)
biplot(wisc.pr)
```

We need a better plot, colored by diagnosis in wisc.df

```{r}
plot(wisc.pr$x[,1], wisc.pr$x[,2], xlab = "PC1", ylab = "PC2", col = wisc.df$diagnosis)
```

Make a scree plot for our PCA
```{r}
wisc.var <- wisc.pr$sdev^2
pve <- round(wisc.var/sum(wisc.var)*100, 2)
plot(pve, type = "b", ylab = "data percentage", xlab = "PC")
barplot(pve, ylab = "Precent of Variance Explained",
     names.arg=paste0("PC",1:length(pve)), las = 2, axes = FALSE)
axis(2, round(pve) )
```

Clustering on the PCA
```{r}
# Distance matrix nested inside hclust function
wisc.pr.hc <- hclust(dist(wisc.pr$x[,1:2]), method = "ward.D2")
plot(wisc.pr.hc)
```

```{r}
grp3 <- cutree(wisc.pr.hc, k = 3)
table(grp3)
```

```{r}
plot(wisc.pr$x[, 1], wisc.pr$x[, 2], xlab = "PC1", ylab = "PC2", col = grp3)
```

```{r}
diagnosis <- wisc.df$diagnosis == "M"
table (grp3, wisc.df$diagnosis)
```

