---
title: "Machine Learning"
author: "Tuan"
date: "2/6/2019"
output: github_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

## R Markdown

k-means clustering

```{r}
tmp <- c(rnorm(30,-3), rnorm(30,3))
x <- cbind(x=tmp, y=rev(tmp))
plot(x)
```

```{r}
km <- kmeans(x, centers = 2, nstart = 20)
km
```

In the result:
*Clustering vector*: a vector shows which element of data belongs to which cluster

Size of cluster
```{r}
km$size
```

Cluster membership vector
```{r}
km$cluster
```

```{r}
plot(x, col = km$cluster)
points(km$centers, col = "blue", pch = 15, cex = 1.5)
```

```{r}
# First we need to calculate point (dis)similarity
#   as the Euclidean distance between observations
dist_matrix <- dist(x)
# The hclust() function returns a hierarchical
#  clustering model
hc <- hclust(d = dist_matrix)
# the print method is not so useful here
hc
# The equivalent in one single line: hclust(dist(x))
```

```{r}
plot(hc)
abline(h = 6, col = "red")
grp2 <- cutree(hc, h = 6)
```
```{r}
plot(x, col = grp2)
```

```{r}
plot(hc)
abline(h = 6, col = "red")
grp6 <- cutree(hc, h = 2.5)
table(grp6)
```
We can also use k = groups to cut tree
```{r}
cutree(hc, k = 3)
```
### Using different Linkage methods of HC
```{r}
d <- dist_matrix
hc.complete <- hclust(d, method="complete")
plot(hc.complete)
hc.average  <- hclust(d, method="average")
plot(hc.average)
hc.single   <- hclust(d, method="single")
plot(hc.single)
```
Made up overlapping data - a bit more real lfe like
```{r}
# Step 1. Generate some example data for clustering
x2 <- rbind(
  matrix(rnorm(100, mean=0, sd = 0.3), ncol = 2),   # c1
  matrix(rnorm(100, mean = 1, sd = 0.3), ncol = 2), # c2
  matrix(c(rnorm(50, mean = 1, sd = 0.3),           # c3
           rnorm(50, mean = 0, sd = 0.3)), ncol = 2))
colnames(x2) <- c("x", "y")
# Step 2. Plot the data without clustering
plot(x2)
# Step 3. Generate colors for known clusters
#         (just so we can compare to hclust results)
col <- as.factor( rep(c("c1","c2","c3"), each=50) )
plot(x2, col=col)
# Use the dist(), hclust(), plot(), cutree(). Hierarchical clustering goes through 3 main steps: distance matrix, hclust, and plot.
dist_matrix2 <- dist(x2)
hclust2 <- hclust(dist_matrix2)
groups3 <- cutree(hclust2, k = 3)
plot(hclust2)
plot(x2, col = groups3)
# Each run of clustering will have slightly different colored graph due to random assignment at the beginning of the algorithm.
```

# Principal component analysis (PCA)
```{r}
mydata <- read.csv("https://tinyurl.com/expression-CSV",
row.names=1)
head(mydata)
```

**NOTE**: **prcomp()** expects the samples to be rows and genes to be columns so we need to first transpose the matrix with the **t()** function!
```{r}
pca <- prcomp(t(mydata), scale = T)
summary (pca)
```

Make our first PCA plot
```{r}
plot(pca$x[, 1], pca$x[, 2], xlab = "PC1", ylab = "PC2")
```

```{r}
## Precent variance is often more informative to look at
pca.var <- pca$sdev^2
pca.var.per <- round(pca.var/sum(pca.var)*100, 1)
 barplot(pca.var.per, main="Scree Plot",
 xlab="Principal Component", ylab="Percent Variation")
```

```{r}
## A vector of colors for wt and ko samples
 colvec <- as.factor( substr( colnames(mydata), 1, 2) )
#substr(x, start, stop): take a string a trim it down from start to stop
plot(pca$x[,1], pca$x[,2], col=colvec, pch=16,
xlab=paste0("PC1 (", pca.var.per[1], "%)"),
ylab=paste0("PC2 (", pca.var.per[2], "%)"))
```

```{r}
UK.food <- read.csv("https://bioboot.github.io/bggn213_f17/class-material/UK_foods.csv", row.names = 1)
head(UK.food)
UK.food.pca <- prcomp(t(UK.food))
plot(UK.food.pca$x[,1], UK.food.pca$x[,2], xlab = "PC1", ylab = "PC2")
text(UK.food.pca$x[,1], UK.food.pca$x[,2], colnames(UK.food), col = c("orange", "red", "blue", "dark green"))
```
```{r}
## Lets focus on PC1 as it accounts for > 90% of variance 
par(mar=c(10, 3, 0.35, 0))
barplot( UK.food.pca$rotation[,1], las=2 )
```

