---
title: "Lecture 12 Drug Discovery with Structural Bioinformatics"
author: "Tuan"
date: "2/20/2019"
output: github_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

## Clean up our protein trarget structure

First we download a target (i.e protein receptor) structure from the main PDB database. We will pick PDB ID "1hsg"
```{r}
library(bio3d)

pdb.code <- "1hsg"
file.name <- get.pdb(pdb.code)
```

Extract the protein only segment of this PDB entry and write out a new PDB format file. We will also do the same thing for the ligand
```{r}
hiv <- read.pdb(file.name)
hiv
```

protein extraction first
```{r}
prot <- trim.pdb(hiv, "protein")
prot
```

```{r}
write.pdb(prot, file = paste(pdb.code, "_protein.pdb", sep = ""))
```

Do the same thing for the ligand
```{r}
ligand <- trim.pdb(hiv, "ligand")
write.pdb(ligand, file = paste(pdb.code, "_ligand.pdb", sep = ""))
```

The last 2 columns of .pdbqt file represent electrostatic charges of atoms.

Convert the docking result for viewing in VMD
```{r}
res <- read.pdb("all.pdbqt", multi = T)
res
```

```{r}
write.pdb(res, file = "result.pdb")
```

```{r}
res <- read.pdb("all.pdbqt", multi=TRUE) 
ori <- read.pdb("ligand.pdbqt")
rmsd(ori, res)
```

```{r}
rmsd(res)
```

